# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
We first read the data into a variable
```{r, echo=TRUE}
activity <- read.csv("activity.csv", header=TRUE, na.strings="NA")
```

After this we need to make sure the date variable is of class Date
```{r, echo=TRUE}
if( class(activity$date) != 'Date'){
  activity$date <- as.Date(activity$date)  
}
```

## What is mean total number of steps taken per day?
We need to get the average steps taken for each day of the week. For this case we use tapply()
function and we need to set na.rm=TRUE so we remove the missing values.
```{r echo=TRUE}
#we use simplify=FALSE so we get a list (with dims) that is easier to deal with than a 
#single dim array
Mean <- tapply(activity$steps, weekdays(activity$date, abbreviate=T), mean, na.rm=TRUE, simplify=FALSE)

#make the list a data frame, that is easy to deal with
means <- as.data.frame(Mean)

#add a new column with the weekdays that we can later use in our manipulations
means$Weekday <- rownames(means)

#sort based on the weekdays, first set the levels
means$Weekday <- factor(means$Weekday, 
                        levels=c("Sun", "Mon","Tue", "Wed", 
                                "Thu", "Fri","Sat"))
sortweek <- means[order(means$Weekday, decreasing=FALSE),]
```

Then we are making the histogram
```{r, echo=TRUE}
#make the Mean variable all numeric
sortweek$Mean <- as.numeric(sortweek$Mean)

#plot
with(sortweek, plot(Weekday, Mean), type="p", main="Daily Number of Steps (Average)",
                    xlab="Week Days", ylab="Average Steps")
```
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?





